---
title: "CustomerPersonalityAnalysis(Statistic)"
author: "Allen Wu"
date: "2023-11-16"
output: html_document
---


# 1.import labarary
```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(car)
```



# 2.import dataset
```{r}
df = read.csv("marketing_campaign.csv", sep = "\t", header=T)
head(df)
```


# 3.Data Cleaning
- Fill in missing values in income by calculating the group mean (by Education).
- Convert the birthdates to ages for presentation in subsequent visual analyses.
- Merge Kidhome and Teenhome into a single feature called Children to reduce the number of required features.
- 

```{r}
# Fill NA on variable "Income"
df <- df %>%
  group_by(Education) %>%
  mutate(Income = ifelse(is.na(Income), mean(Income, na.rm = TRUE), Income)) %>%
  ungroup()

# Convert 'Year_Birth' to 'Age' for better understanding
df$Age <- 2024 - df$Year_Birth
df <- df[, !(names(df) %in% "Year_Birth")]  # Remove the original 'Year_Birth' column


# "Education" variable
#df$Education <- recode(df$Education,
                       #"Basic" = "Undergraduate",
                       #"2n Cycle" = "Undergraduate",
                       #"Graduation" = "Graduate",
                       #"Master" = "Postgraduate",
                       #"PhD" = "Postgraduate")

# Merge 'Kidhome' and 'Teenhome' into 'Children'
df$Children <- df$Kidhome + df$Teenhome
df <- df[, !(names(df) %in% c("Kidhome", "Teenhome"))]  # Remove the original 'Kidhome' and 'Teenhome' columns

# Referring to the parenthood status
df$Parental_Status <- as.integer(df$Children != 0)

# Count Customer's total spent
df$Total_Spent <- df$MntWines + df$MntFruits + df$MntMeatProducts +
                  df$MntFishProducts + df$MntSweetProducts + df$MntGoldProds

# Count total Cmp accepted by Customer
df$Total_Offer <- df$AcceptedCmp1 + df$AcceptedCmp2 + df$AcceptedCmp3 +
                  df$AcceptedCmp4 + df$AcceptedCmp5

# Count total purchases by all methods
df$Num_Total_Purchases <- df$NumWebPurchases + df$NumCatalogPurchases +
                          df$NumStorePurchases + df$NumDealsPurchases
head(df)
```

# 4.Descriptive Statistic
```{r}
# numerical variables
print('Basic Statistic for numurical variables')
summary(df)

# categorical variables
print("Basic Statistic for categorical variables")
table_counts <- table(df$Education)
percentage_counts <- prop.table(table_counts) * 100
summary_table <- data.frame(Counts = table_counts, Percentages = percentage_counts)
print(summary_table)

table_counts <- table(df$Marital_Status)
percentage_counts <- prop.table(table_counts) * 100
summary_table <- data.frame(Counts = table_counts, Percentages = percentage_counts)
print(summary_table)

```
```{r}
# Plotting Distribution

# Specify numeric features
numeric_features <- c('Total_Spent', 'Num_Total_Purchases', 'Age', 'Income', 'Recency', 'MntWines', 'MntFruits', 'MntMeatProducts', 'MntFishProducts', 'MntSweetProducts', 'MntGoldProds')

# Loop through numeric features to create histograms
for (feature in numeric_features) {
  # Set up the plot
  plt <- ggplot(df, aes(x = .data[[feature]])) +
    geom_histogram(binwidth = (max(df[[feature]] ) - min(df[[feature]])) / 30, fill = "skyblue", color = "black", alpha = 0.7) +
    labs(title = paste("Distribution of", feature),
         x = feature,
         y = "Frequency") +
    theme_minimal()

  # Show the plot
  print(plt)
}

# Specify categorical features
categorical_features <- c('Education', 'Marital_Status')

# Loop through categorical features to create count plots
for (feature in categorical_features) {
  # Set up the plot
  plt <- ggplot(df, aes(x = .data[[feature]])) +
    geom_bar(fill = "skyblue", color = "black", alpha = 0.7) +
    labs(title = paste("Count of customers by", feature),
         x = feature,
         y = "Count") +
    theme_minimal()

  # Show the plot
  print(plt)
}
```

# Customer Segmentation Analysis
```{r}
# Explore relationship

# Scatter Plot
# Select the numeric features
numeric_features <- c('Total_Spent', 'Num_Total_Purchases', 'Age', 'Income', 'Recency', 'MntWines', 'MntFruits', 'MntMeatProducts', 'MntFishProducts', 'MntSweetProducts', 'MntGoldProds')

# Create scatter plots for each numeric feature against Total_Spent
for (feature in numeric_features) {
  scatter_plot <- ggplot(df, aes(x = df[[feature]], y = Total_Spent)) +
    geom_point() +
    labs(title = paste("Scatter Plot between", feature, "and Total_Spent"),
         x = feature,
         y = "Total_Spent") +
    theme_minimal()

  print(scatter_plot)
}
  

```


```{r}
# Analyze "Income" on diff "Education"
# Set up the plot with group means
plt <- ggplot(df, aes(x = Education, y = Income, fill = Marital_Status)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  labs(title = "Mean Income Across Demographic Groups",
       x = "Education",
       y = "Mean Income",
       fill = "Marital Status") +
  theme_minimal()

# Show the plot
print(plt)

# Analyze "Total Spend" on diff "Education"
# Set up the plot with group means
plt <- ggplot(df, aes(x = Education, y = Total_Spent, fill = Marital_Status)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  labs(title = "Mean Income Across Demographic Groups",
       x = "Education",
       y = "Mean Total_Spent",
       fill = "Marital Status") +
  theme_minimal()

# Show the plot
print(plt)

```
# Purchasing Behavior Analysis
```{r}
# Specify the product categories
product_categories <- c('MntWines', 'MntFruits', 'MntMeatProducts', 'MntFishProducts', 'MntSweetProducts', 'MntGoldProds')

# Create pie charts for each product category
# Calculate percentage of spending for each product category
percentage_spending <- colMeans(df[, product_categories]) / sum(colMeans(df[, product_categories])) * 100

# Create a data frame for the pie chart
pie_data <- data.frame(category = names(percentage_spending), percentage = percentage_spending)

# Create a pie chart
pie_chart <- ggplot(pie_data, aes(x = "", y = percentage, fill = category)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y") +
  labs(title = "Percentage of Spending on Different Product Categories",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Show the pie chart
print(pie_chart)

# Calculate percentage of spending for each product category by Marital_Status
percentage_spending <- aggregate(df[, product_categories], by = list(df$Marital_Status), FUN = function(x) sum(x) / sum(df[, product_categories]) * 100)

# Reshape the data for plotting
percentage_spending_long <- tidyr::gather(percentage_spending, key = "Product_Category", value = "Percentage", -Group.1)

# Create a bar plot
bar_plot <- ggplot(percentage_spending_long, aes(x = Product_Category, y = Percentage, fill = factor(Group.1))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage of Spending on Different Product Categories by Marital Status",
       x = "Product Category",
       y = "Percentage",
       fill = "Marital Status") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(bar_plot)


# Calculate percentage of spending for each product category by "Education"
percentage_spending <- aggregate(df[, product_categories], by = list(df$Education), FUN = function(x) sum(x) / sum(df[, product_categories]) * 100)

# Reshape the data for plotting
percentage_spending_long <- tidyr::gather(percentage_spending, key = "Product_Category", value = "Percentage", -Group.1)

# Create a bar plot
bar_plot <- ggplot(percentage_spending_long, aes(x = Product_Category, y = Percentage, fill = factor(Group.1))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage of Spending on Different Product Categories by Education",
       x = "Product Category",
       y = "Percentage",
       fill = "Education") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(bar_plot)



```
```{r}

# Box plot
box_plot <- ggplot(df, aes(x = as.factor(Children), y = Total_Spent)) +
  geom_boxplot(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Total Spending by Number of Children/Teenagers",
       x = "Number of Children/Teenagers",
       y = "Total Spending") +
  theme_minimal()

# Show the plots
print(box_plot)


```

# Web and Store Purchases Analysis
```{r}
# Create box plots for each purchase channel
purchase_channels <- c('NumWebPurchases', 'NumCatalogPurchases', 'NumStorePurchases')

# Create box plots
box_plots <- list()

for (channel in purchase_channels) {
  box_plot <- ggplot(df, aes(x = as.factor(channel), y = df[[channel]])) +
    geom_boxplot(fill = "skyblue", color = "black", alpha = 0.7) +
    labs(title = paste("Distribution of Purchases through", channel),
         x = "Purchase Channel",
         y = "Number of Purchases") +
    theme_minimal()

  box_plots[[channel]] <- box_plot
}

# Print the box plots
for (channel in purchase_channels) {
  print(box_plots[[channel]])
}

# Plotting box plot for purchase chanel by Marital_Status
for (channel in purchase_channels) {
  box_plot <- ggplot(df, aes(x = Marital_Status, y = df[[channel]])) +
    geom_boxplot(fill = "skyblue", color = "black", alpha = 0.7) +
    labs(title = paste("Distribution of", channel, "by Marital_Status"),
         x = "Marital_Status",
         y = paste("Number of", channel)) +
    theme_minimal()

  # Show the box plot for each channel
  print(box_plot)
}
```

# Discount Analysis
```{r}
# Analyze the distribution of deals purchases
# Create a box plot for deals purchases by Marital_Status
box_plot <- ggplot(df, aes(x = Marital_Status, y = NumDealsPurchases)) +
  geom_boxplot(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Deals Purchases by Marital_Status",
       x = "Marital_Status",
       y = "Number of Deals Purchases") +
  theme_minimal()

# Show the box plot
print(box_plot)
```
# Hypothesis Testing
```{r}
# Hypothesis: Income Affects Spending

# Create a new variable 'IncomeGroup' based on income levels
customer_data <- df %>%
  mutate(IncomeGroup = cut(Income, breaks = c(0, 30000, 60000, 90000, 120000, Inf),
                           labels = c("0-30k", "30k-60k", "60k-90k", "90k-120k", "120k+"),
                           include.lowest = TRUE))

# Check the summary statistics for the spending variables by income group
summary_table <- customer_data %>%
  group_by(IncomeGroup) %>%
  summarise(
    Mean_Wines = mean(MntWines),
    Mean_Fruits = mean(MntFruits),
    Mean_MeatProducts = mean(MntMeatProducts),
    Mean_FishProducts = mean(MntFishProducts),
    Mean_SweetProducts = mean(MntSweetProducts),
    Mean_GoldProds = mean(MntGoldProds)
  )

# Print summary table
print(summary_table)

# Perform ANOVA test for each spending variable
# Example for 'Wines' spending
anova_result_wines <- aov(MntWines ~ IncomeGroup, data = customer_data)
print(summary(anova_result_wines))

# Repeat the ANOVA test for other spending variables
# Example for 'Fruits' spending
anova_result_fruits <- aov(MntFruits ~ IncomeGroup, data = customer_data)
print(summary(anova_result_fruits))

# Repeat for other spending variables

# Post-hoc pairwise t-tests to identify specific income groups that differ significantly
# Example for 'Wines' spending
posthoc_wines <- TukeyHSD(anova_result_wines)
print(posthoc_wines)

# Repeat post-hoc tests for other spending variables
# Example for 'Fruits' spending
posthoc_fruits <- TukeyHSD(anova_result_fruits)
print(posthoc_fruits)

```
```{r}
# Hypothesis: Education Influences Total Spending
# Assuming your dataframe is named 'customer_data'

# Create a new variable 'EducationGroup' based on education levels
customer_data <- customer_data %>%
  mutate(EducationGroup = as.factor(Education))

# Check unique values in the 'Education' column and recode if necessary
unique(customer_data$Education)

# Recode if needed (replace 'Basic' with a specific level)
# customer_data$Education <- recode(customer_data$Education, 'Basic' = 'Basic_Level')

# Check the summary statistics for the spending variables by education group
summary_table_education <- customer_data %>%
  group_by(EducationGroup) %>%
  summarise(
    Total_Spending = MntWines + MntFruits + MntMeatProducts + MntFishProducts + MntSweetProducts + MntGoldProds
  )

# Print summary table
print(summary_table_education)

# Perform ANOVA test for total spending
anova_result_education <- aov(Total_Spent ~ EducationGroup, data = customer_data)
print(summary(anova_result_education))

# Post-hoc pairwise t-tests to identify specific education groups that differ significantly
posthoc_education <- TukeyHSD(anova_result_education)
print(posthoc_education)

```
```{r}
# Hypothesis: Parenthood Affects Total Spending
# Assuming your dataframe is named 'customer_data'

# Create a new binary variable 'HasChildren' indicating whether the customer has children
customer_data <- df %>%
  mutate(HasChildren = ifelse(Children > 0, "With Children", "Without Children"))

# Check the summary statistics for the spending variables by parenthood status
summary_table_parenthood <- customer_data %>%
  group_by(HasChildren) %>%
  summarise(
    Total_Spending = MntWines + MntFruits + MntMeatProducts + MntFishProducts + MntSweetProducts + MntGoldProds
  )

# Perform t-test for total spending between customers with and without children
t_test_result <- t.test(Total_Spent ~ Parental_Status, data = customer_data)
print(t_test_result)

```
```{r}
# Hypothesis: Age and Spending Habits
# Check the correlation between age and total spending
correlation_result <- cor.test(df$Age, 
                               df$Total_Spent, 
                               method = "pearson")

# Print correlation result
print(correlation_result)

```
```{r}
# Hypothesis: Web Visits Influence Web Purchases
# Check the correlation between web visits and web purchases
correlation_result <- cor.test(df$NumWebVisitsMonth, 
                               df$NumWebPurchases, 
                               method = "pearson")

# Print correlation result
print(correlation_result)

# Perform linear regression
regression_model <- lm(NumWebPurchases ~ NumWebVisitsMonth, data = customer_data)

# Print regression summary
summary(regression_model)
```

